### 2021-06-10 outlook.tex
Add labels to sec's, fix adem/cartan claims, add keyword: operads

\section{Outlook} \label{s:outlook}

This article looked at Steenrod operations from an algebraic viewpoint, a subject with rich geometric and combinatorial components as well. For example, \cite{Postnikov} and \cite{medina2018prop2} independently introduced equivalent geometric representations of the cup-$i$ products in terms of stabilized arc surfaces \cite{KLP} and weighted ribbon graphs respectively. In fact, an entire $E_\infty$-operad (PROP) is constructed geometrically in this way. We can also interpret the May-Steenrod structure in $U(\mathcal M)$ from the ``oriented surface" perspective, where the norm map $N$ of cyclic groups ---a key ingredient in Definition \ref{def: minimal cyclic resolution}--- was identified in \cite{KLP} with a Dehn twist operator in connection with Connes' cyclic complex.

In Higher Category Theory, the paper \cite{medina2020globular} constructs a functor producing strict \mbox{$\infty$-cat}egories from group-like cup-$i$ coalgebras in a manner similar to \cite{steiner2004omega}. In particular, the cup-$i$ constructions described in this article for standard simplices and cubes define, respectively, the Street nerve and cubical nerve of strict $\infty$-categories. We anticipate that the more general Steenrod $(p,i)$-products constructed in this work will also have deep combinatorial interpretations.

There is a functorial approach to the theory using the formalism of Feynman categories \cite{feynman},
which renders all the structures and notions natural. This includes cyclic, planar cyclic as well as Berger's pre-operads \cite{BergerRecog}. Their interplay is of independent interest \cite{BergerKaufmann,feyrep} and will be linked directly to the constructions of this paper.

In this article we have not focused on the operations that exist non-trivially for \mbox{$E_n$-algebras} with $n$ finite, see part III of \cite{may76homology}. A treatement close to ours for the $E_2$ case was given in \cite{Tourtchine}. Effective constructions for these Dyer-Lashof-Cohen operations should be related to the geometry and combinatorial structure of configuration spaces \cite{KZhang,sinha2013littledisks,berger04combinatorial,ayala2014configuration} and higher categories \cite{Bathigher, BalFiedSchwVogt, Rezkhigher}, and will appear elsewhere.