### 2021-08-20 may_st.tex

\documentclass{amsart}
\include{aux/style}
\include{aux/usualcmds}

\title{Cochain level May--Steenrod operations}

\author[R. Kaufmann]{Ralph M.\ Kaufmann}
\address{R. Kaufmann, Department of Mathematics, Department of Physics and Astronomy, Purdue University}
\email{rkaufman@purdue.edu}

\author[A. Medina-Mardones]{Anibal M. Medina-Mardones}
\address{A. Medina-Mardones, Max Planck Institute for Mathematics \and
	University of Notre Dame du Lac}
\email{ammedmar@mpim-bonn.mpg.de}

\subjclass[2010]{55S05, 55U05, 55S10, 55S12; 55-04, 55U15, 55S15}

\keywords{Steenrod operations, Dyer--Lashof operations, operads, simplicial cochains, cubical cochains}

\include{aux/figures}

\begin{document}
	\input{aux/abstract}
	\maketitle
	\input{sec/introduction}
	\input{sec/preliminaries}
	\input{sec/groups}
	\input{sec/operads}
	\input{sec/steenrod}
	\input{sec/effective}
	\input{sec/outlook}
	\printbibliography
\end{document}