### 2021-08-28 may_st.tex
Add hyperref to email addresses

\documentclass{amsart}
\include{aux/style}
\include{aux/usualcmds}

\title{Cochain level May--Steenrod operations}

\author[R.~Kaufmann]{Ralph~M.~Kaufmann}
\address{R.K., Department of Mathematics, Department of Physics and Astronomy, Purdue University}
\email{\href{mailto:rkaufman@purdue.edu}{rkaufman@purdue.edu}}

\author{Anibal~M.~Medina-Mardones}
\address{A.M-M., Max Plank Institute for Mathematics \and University of Notre Dame}
\email{\href{mailto:ammedmar@mpim-bonn.mpg.de}{ammedmar@mpim-bonn.mpg.de}}

\subjclass[2010]{55S05, 55U05, 55S10, 55S12; 55-04, 55U15, 55S15}

\keywords{Steenrod operations, Dyer--Lashof operations, operads, simplicial cochains, cubical cochains}

\include{aux/figures}

\begin{document}
	\input{aux/abstract}
	\maketitle
	\input{sec/introduction}
	\input{aux/acknowledgement}
	\input{sec/preliminaries}
	\input{sec/groups}
	\input{sec/operads}
	\input{sec/steenrod}
	\input{sec/effective}
	\input{sec/outlook}
	\printbibliography
\end{document}